---
title: "Jackson Mouse Study"
author: "Philip LoGerfo"
date: "May 2021"
output: html_document

---

##########################################
# Program to summarize data from a directory containing individual CSV files and 
# corresponding PNG files into one data file. 
# The program calculates void volumes, measured in microliters, by standardizing
# the void area sizes, measured in pixel counts, to the size of the image of the filter
# paper, and then applying the formula to convert area size to volume size. 
# The void area sizes are contained in CSV files and the images of the filters are in 
# PNG files. 
# For each CSV file and its corresponding PNG file, the following values are returned: 
# Total Volume, Number of Micro Voids, Number of Primary Voids, Mean Total Void Volume 
# Mean Micro Void Volume, Mean Primary Void Volume. These values are added to a data frame,
# and the corresponding CSV file names are used as the rownames for the data frame.   
#########################################

```{r get dependencies}
  
library(png)
library(tidyverse)
 
```


#############

```{r read in csv and png files}

 # !!Make sure to check the name of the directory path!!
 #pngPathName <- "~/Test_set_Lg/Test_Set_Lg"

 trunk <- "~/150 Weeks"

 # Make a list of the PNG file names with the paths included
 pngPaths <- dir(path =trunk, pattern = "*.png", full.names = TRUE, recursive = TRUE)
 
 # Make a list of the PNG file names without the paths
 #pngNames <- dir(path =trunk, pattern = "*.png", full.names = FALSE, recursive = TRUE)

# !!Make sure to check the name of the directory path!!
# csvPathName <-	"~/Test_set_Lg/Test_Set_Lg"
 
 # Make a list of the CSV file names with the paths included
 #csvPaths <- dir(path =trunk, pattern = "*.csv", full.names = TRUE, recursive = TRUE)
 
 csvPaths <- dir(path =trunk, pattern = "*areas.csv", full.names = TRUE, recursive = TRUE)
 # Make a list of the CSV file names without the paths
 #csvNames <- dir(path =trunk, pattern = "*areas.csv", full.names = FALSE, recursive = TRUE)



```

#################


```{r warning = FALSE }

##Convert pixel counts to area (mm^2) and then convert void area to volumes

# Below are a bunch of variables that are initialized before the operation

age <- NULL
diet <- NULL
group <- NULL
day <- NULL
imageNum <- NULL

pixelCount <- NULL
totalVol <- NULL
microVoid <- NULL
primeVoid <- NULL
totalVoids <- NULL
primeAvg = NULL
microAvg = NULL
totalAvg <- NULL


if (length(csvPaths) != length(pngPaths)) {
  print("Number of csv files: ", length(csvPaths), ". Number of png files: ", length(pngPaths), 
        ". Must have equal number of files!")
} else {

  
  for (i in 1:length(csvPaths)) {

	
    branches <- str_split(csvPaths[i],"/")
  
    age <- rbind(age, branches[[1]][5])
    diet <- rbind(diet, branches[[1]][6])
    group <- rbind(group, branches[[1]][7])
    day <- rbind(day, branches[[1]][8])
    imageNum <- rbind(imageNum, branches[[1]][10])
  
    # Read in the PNG file that corresponds with the CSV file
    pp <- readPNG(pngPaths[i])
    # estimate the pixel area of the cage floor
    temppixelCount <- length(which(pp > .05))
    pixelCount <- c(pixelCount, temppixelCount)
  
    # Read in the CSV file 
    pf <- read.delim(csvPaths[i], sep = ",", header = TRUE, row.names = 1)
  
    # make sure there are pee events
    if (length(pf[,1]) > 0) {
    
      #convert pixels to area
      stdArea <- temppixelCount/(265*111)
    
      #calculate volume 
      pvol <- (pf[1]/stdArea)*0.283
    
      # filter out very small volumes
      filter <- which(pvol > 2)
    
      # make a list of voids for file i
      fpf <- pvol[filter,1]
    
      # check again for pee events after filtering
      if (length(fpf) > 0) {
      
        # count the number of total voids
        tempEvent <- length(fpf)
        #add the total void count from that file to list totalVoids
        totalVoids <- c(totalVoids, tempEvent)
      
        # count microvoids
        tempMicro <- which(fpf < 22)
        if (length(tempMicro > 0)) {
          # add the total micro void count to list microv
          microVoid <- c(microVoid, length(tempMicro))
          # makes a list of micro voids only
          mfpf <- fpf[tempMicro]
          # calculate the mean of micro voids
          tempMicro <- mean(mfpf)
          # add the mean of micro voids to list microAvg
          microAvg <- c(microAvg, tempMicro)
        } else {
            # if there are no micro voids
            microVoid <- c(microVoid, 0)
            microAvg = c(microAvg, 0)
          }
      
        # count primaryvoids
        tempPrime <- which(fpf >= 22)
        if (length(tempPrime > 0)) {
          # add the total primary void count to list primeVoid
          primeVoid <- c(primeVoid, length(tempPrime))
          # makes a list of primary voids only
          pfpf <- fpf[tempPrime]
          # calculate the mean of primary voids
          tempPrimeAvg <- mean(pfpf)
          # add the mean of primary voids to list primeAvg
          primeAvg <- c(primeAvg, tempPrimeAvg)
        
        } else {
            # if there are no primary voids
            primeVoid <- c(primeVoid, 0)
            primeAvg = c(primeAvg, 0)
          }
      
        # sum the total area of all pee events   
        tempa <- sum(fpf)
        totalVol <- c(totalVol, tempa)
      
        # calculate the mean of all voids
        tempAvg <- mean(fpf)
        totalAvg <- c(totalAvg, tempAvg)
      } else {
          # If there are no micro or primary voids, but some voids < 1500 in area
          microVoid <- c(microVoid, 0)
          primeVoid <- c(primeVoid, 0)
          primeAvg = c(primeAvg, 0)
          microAvg = c(microAvg, 0) 
          totalVol <- c(totalVol, 0)
          totalVoids <- c(totalVoids,  0)
          totalAvg <- c(totalAvg, 0)
        }
    
    } else {
    
        totalVol <- c(totalVol, 0)
        totalVoids <- c(totalVoids, 0)
        microVoid <- c(microVoid, 0)
        primeVoid <- c(primeVoid, 0)
        totalAvg <- c(totalAvg, 0)
        primeAvg = c(primeAvg, 0)
        microAvg = c(microAvg, 0)
      } 
  
    if (i %% 5 == 0)
      print(paste(i, " of ", length(csvPaths), sep = ""))
  
  }



# Format the results into a table and write to csv file 

  pdata <- data.frame(row.names = imageNum, Image = imageNum, Total_Volumes = totalVol, Total_Voids = totalVoids, 
                    Total_Mean = totalAvg, Primary_Voids = primeVoid, Primary_Mean = primeAvg, 
                    Micro_Voids = microVoid, Micro_Mean = microAvg, Age = age, Diet = diet, Group = group, Day = day) 


  # !!Make sure to check the name of the file that is written!!
  nameOfFile <- "VoidDataSummary.csv"

  write.table(pdata, nameOfFile, sep = ",", row.names = FALSE)

}


######### END ############### END ################## END ################ END ############

```
###################

Look at the nameOfFile variable. That is the name of the file that will be written with all the results in it.
You can change it to anything you want. Make sure to put the name in quotes "".


```{r }

```

